Lattices of acyclic pipe dreams

### 2023-01-20 acyclicFacets/draft/latticePipeDreams.tex
wrapping up

@@

% theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{theoremA}{Theorem}
\renewcommand{\thetheoremA}{\Alph{theoremA}}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{conjectureA}{Conjecture}
\renewcommand{\theconjectureA}{\Alph{conjectureA}}
\crefname{conjecture}{Conjecture}{Conjectures}

@@

\vspace*{.2cm}
\maketitle

@@

@@

This paper develops a similar framework for acyclic pipe dreams.
Pipe dreams were introduced by N.~Bergeron and S.~Billey in~\cite{BergeronBilley} to compute Schubert polynomials and later revisited in the context of Gr\"obner geometry by A.~Knutson and E.~Miller~\cite{KnutsonMiller-GroebnerGeometry}, who coined the name \emph{pipe dreams}.
They have important connections and applications to various areas related to Schubert calculus and Schubert varieties~\cite{LascouxSchutzenberger-PolynomesSchubert, LascouxSchutzenberger-SchubertLittlewoudRichardson}. 
A pipe dream is an arrangement of pipes in the triangular shape, each entering along the vertical side and exiting along the horizontal side.
They are grouped according to their exiting permutation, given by the order in which the pipes appear along the horizontal axis.
The linear extensions of a pipe dream are the permutations of its pipes such that for each contact the northwest pipe appears before the southeast pipe in the permutation.
The pipe dreams with at least one linear extension are called acyclic and naturally appear in the study of brick polytopes~\cite{PilaudSantos-brickPolytope}.
A flip in a pipe dream exchanges a contact with a crossing between two pipes, and the flip is increasing when the contact is southwest of the crossing involved in the flip.
A brief recollection on pipe dreams is given in \cref{sec:preliminaries}.

@@

\begin{theoremA}
For any permutation~$\omega$, the Hasse diagram of the lattice quotient of the interval~$[e,\omega]$ of the weak order by the pipe dream congruence of~$\omega$ is isomorphic to the increasing flip graph on acyclic pipe dreams with exiting permutation~$\omega$.
\end{theoremA}

@@

We then explore in \cref{sec:furtherTopics} some natural further topics.
We describe two algorithms to compute the unique acyclic pipe dream whose linear extensions contain a given permutation generalizing the binary tree insertion map on permutations (\cref{subsec:sweepingAlgorithm,subsec:insertionAlgorithm}), we describe the pipe dream congruence as the transitive closure of a local rewriting rule generalizing that of the sylvester congruence (\cref{subsec:rewritingRule}), and we describe a natural commutative diagram of lattice morphism generalizing the connection between the recoil map and the binary tree insertion map on permutations and the canopy on binary trees (\cref{subsec:canopy}).

@@

\begin{theoremA}
For any non-empty subword complex~$\subwordComplex(Q, \omega)$, the sets~$\linearExtensions(I)$ for all facets~$I$ of~$\subwordComplex(Q, \omega)$ are order convex and form a partition of a lower set of the weak order that contains the interval~$[e, \omega]$.
\end{theoremA}

@@

\begin{theoremA}
If~$Q$ is sorting, then the sets~$\linearExtensions(I)$ for all facets~$I$ of~$\subwordComplex(Q, \omega)$ form a partition of the interval~$[e, \omega]$.
\end{theoremA}

In the case when~$[e, \omega]$ does not contain all sets of linear extensions, it is natural to consider the restriction of this partition to~$[e, \omega]$ by the sets~$[e, \omega] \cap \linearExtensions(\omega)$.
This defines an equivalence relation~$\equiv_{Q, \omega}$ on~$[e, \omega]$ that we call the subword complex congruence.
However, as the sets of linear extensions are not always intervals of the weak order, this equivalence relation is not always a lattice congruence.
This can seems to be fixed by an additional assumption on~$Q$.
We say that~$Q$ is alternating when all non-commuting pairs $s, t\in S$ alternate within $Q$ (this notion was already considered in \cite{PilaudSantos-brickPolytope, CeballosLabbeStump}).

\begin{conjectureA}
If~$Q$ is alternating, then the subword complex congruence~$\equiv_{Q, \omega}$ is a lattice congruence of the interval~$[e, \omega]$ of the weak order.
\end{conjectureA}

In contrast to the case of pipe dreams, the Hasse diagram of the quotient~$[e, \omega]/{\equiv_{Q, \omega}}$ is not isomorphic to the increasing flip graph on acyclic facets of~$\subwordComplex(Q, \omega)$ for two reasons:
\begin{itemize}
\item First, not all acyclic facets of~$\subwordComplex(Q, \omega)$ appear as vertices of~$[e, \omega]/{\equiv_{Q, \omega}}$ (see \cref{rem:linearExtensionsPartitionSubwordComplexA}). We say that a facet is strongly acyclic if~$[e, \omega] \cap \linearExtensions(\omega) \ne \varnothing$.
\item Second, not all flips between two strongly acyclic facets of~$\subwordComplex(Q, \omega)$ define a cover relation of~$[e, \omega]/{\equiv_{Q, \omega}}$ (see \cref{exm:acyclicFlipvsWeakOrder}). We say that the flip of a position~$i$ in a facet~$I$ is external if the root~$\rootFunction{I}{i}$ directing the flip is a ray of the root configuration~$\Roots(I)$.
\end{itemize}
This leads us to the following conjecture.

\begin{conjectureA}
If~$Q$ is alternating, then the Hasse diagram of the quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between strongly acyclic facets of~$\subwordAcyclicFacets(Q, \omega)$.
\end{conjectureA}

Combining the sorting and alternating conditions, we thus obtain the following conjecture, which can be seen as the natural extension of our pipe dream result.

\begin{conjectureA}
If~$Q$ is sorting and alternating, then the Hasse diagram of the quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between acyclic facets of~$\subwordAcyclicFacets(Q, \omega)$.
\end{conjectureA}

This last conjecture has a strong connection to the geometry of subword complexes given by brick polytopes of~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and the brick polyhedra of~\cite{JahnStump}.
Namely, the extremal increasing flips between acyclic facets of~$\subwordAcyclicFacets(Q, \omega)$ is precisely the bounded part (meaning forgetting the unbounded rays) of the brick polyhedron~$\brickPolyhedron(Q, \omega)$ oriented by a natural linear functional.
In particular, when~$\omega = \wo$, we obtain the following conjecture, extending results from~\cite{Pilaud-brickAlgebra}.

\begin{conjectureA}
If~$Q$ is sorting and alternating, the oriented graph of the brick polytope~$\brickPolyhedron(Q, \wo)$ is the Hasse diagram of a lattice quotient of the weak order.
\end{conjectureA}

@@

\begin{figure}[ht]
	\centerline{
		%\input{pipeDreams}
		\includegraphics[scale=.9]{pipeDreams}
	}
	\caption{Two pipe dreams of~$\pipeDreams(1365724)$ connected by a flip (exchanging a contact with the crossing on the two red pipes~$3$ and~$4$).}
	\label{fig:pipeDreams}
\end{figure}

@@

\begin{example}
\label{exm:Tamari1}
We say that a pipe dream is \defn{reversing} if it fixes the first and last pipes and reverses the order of the remaining pipes.
In this case, it is natural to label the pipes from~$0$ to~$n+1$, so that the permutation of the pipes is~$\rho_n \eqdef [0, n, n-1, \dots, 2, 1, n+1]$.
% We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed in~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the family of reversing pipe dreams belong to the Catalan families, meaning that it is counted by the famous Catalan numbers.
\cref{fig:bijection} illustrates explicit bijections between reversing pipe dreams of~$\pipeDreams(\rho_n)$, binary trees with $n$ internal nodes, and the triangulations of a convex $(n+2)$-gon.
More precisely, the map which sends a contact in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence: \\[.3cm]
\centerline{
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
pipe $i$ of~$P$ & triangle $i\duality$ of~$P\duality$ (with central vertex~$i$), \\
contact between pipes~$i$ and~$j$ of~$P$ & common side of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
crossing between pipes~$i$ and~$j$ of~$P$ & common bisector of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
contact flips in~$P$ & diagonal flips in~$P\duality$. \\
%increasing flip poset & Tamari lattice.
\end{tabular}} \\[.3cm]
%
\begin{figure}
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}
%
As a consequence, this map sends the increasing flip poset on reversing pipe dreams to the \defn{Tamari lattice} on binary trees.
This lattice is defined as the transitive closure of right rotations on binary trees, and is obtained as the quotient of the weak order by the \defn{sylvester congruence}.
\vincent{Not sure where we want that...}
An important point here is that all reversing pipe dreams are acyclic, since there contact graphs are (oriented) binary trees.
\end{example}

@@

\section{Further topics}
\label{sec:furtherTopics}

@@

Our first algorithm to construct~$\insertion{\pi}{\omega}$ is a \defn{sweeping algorithm}, inspired by the algorithm to compute greedy pipe dreams~\cite{PilaudPocchiola, PilaudStump-ELlabelings}.
Extensions of this algorithm will be discussed in \cref{..}. \vincent{todo. Was this algorithm explicitly somewhere before? We should cite Jahn-Stump later in the paper.}

@@

\begin{example}
\label{exm:sweepingAlgorithm}
\cref{fig:sweepingAlgorithm} illustrates the sweeping algorithm for the permutations~${\pi = 513264}$ and~${\omega = 561324}$.
The algorithm has 15 steps, but we have grouped together steps $2$ to $9$ (second arrow) and steps $12$ to $13$ (fifth arrow) as they have the same justification.
We place contacts in steps $1$, $10$, $12$, $13$ and~$15$ since~$\omega^{-1}(i) < \omega^{-1}(j)$, and in step~$14$ since~$i > j$.
We place crossings in steps $2$ to $9$ since~$i < j$, $\omega^{-1}(i) > \omega^{-1}(j)$ and we are in the column~$\omega^{-1}(j)$, and in step $11$ since~$i < j$, $\omega^{-1}(i) > \omega^{-1}(j)$ and~$\pi^{-1}(i) > \pi^{-1}(j)$.
%
\begin{figure}[h]
	\centerline{\includegraphics[scale=.9]{sweeping}}
	\caption{Sweeping algorithm for the permutations~$\pi = 513264$ and~$\omega = 561324$.} % See details in \cref{exm:sweepingAlgorithm}.}
	\label{fig:sweepingAlgorithm}
\end{figure}
\end{example}

@@

\begin{example}
\label{exm:insertionAlgorithm}
\cref{fig:insertionAlgorithm} illustrates the insertion algorithm for the permutations~${\pi = 513264}$ and~${\omega = 561324}$.
For instance, when inserting the (green) pipe~$4$ in the last step, the currently free southeast elbows are the first southeast elbow of the (blue) pipe~$3$ and the two southeast elbows of the (purple) pipe~$2$.
%
\begin{figure}[t]
	\centerline{\includegraphics[scale=.9]{insertion}}
	\caption{Insertion algorithm for the permutations~$\pi = 513264$ and~$\omega = 561324$.}
	\label{fig:insertionAlgorithm}
\end{figure}
\end{example}

@@

The objective of this section is to partially extend our results about the lattice structure of acyclic pipe dreams (\cref{sec:latticeAcyclicPipeDreams}) to the wider context of subword complexes in finite Coxeter groups.
We start with some basic preliminaries on finite Coxeter groups (\cref{subsec:finiteCoxeterGroups}) and subword complexes (\cref{subsec:subwordComplexes}).
We then define linear extensions of facets of subword complexes (\cref{subsec:linearExtensionsFacets}) and present two theorems (\cref{subsec:twoTheorems}) and two conjectures (\cref{subsec:twoConjectures}) about them.
To conclude, we present a sweeping algorithm to construct the facet with a given linear extension (\cref{subsec:sweepingAlgorithmSubwordComplexes})

@@

We now introduce the analogue of \cref{def:linearExtensions} for subword complexes.
\vincent{Is linear extension a good term? Should we say $W$-extension?}

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:cover} and~\eqref{item:convex}.
\begin{enumerate}
\setcounter{enumi}{1}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. \cesarm{add discussion}
%
\begin{figure}[h]
\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
\includegraphics[width=0.43\textwidth]{interval_not_contained2}
\end{figure}
%
\setcounter{enumi}{3}
\item $\linearExtensions(I)$ is not an interval in general. See for instance the example~\cite[Figure~9]{PilaudStump-brickPolytope}. 
In that figure, the elements of the week order $\mathfrak{S}_4$ that are linear extensions of the same acyclic facet of a given subword complex are grouped together. Some of the groups in the figure are not intervals.  \\
\end{enumerate}
\end{remark}

@@

\subsection{Two conjectures on linear extensions of facets}
\label{subsec:twoConjectures}

@@

\begin{example}
\label{exm:acyclicFlipvsWeakOrder}
Consider the type~$A_2$ Coxeter system, the word~$Q = \tau_1 \tau_2 \tau_1 \tau_2 \tau_1 \tau_2$ and the longest element~$\wo = \tau_1 \tau_2 \tau_1 = \tau_2 \tau_1 \tau_2$.

\vincent{work this out}
    Let $W=S_3$ be the symmetric group generated by simple transpositions $S=\{s_1,s_2\}$ with $s_i=(i, i+1 )$.
Let $Q=(s_1,s_2,s_1,s_2,s_1,s_2)$ and $\omega=\wo=s_1s_2s_1=s_2s_1s_2$.
The subword complex $\subwordComplex(Q,\omega)$ has eight facets, six of which are acyclic.
    All the fibers of $\equiv_{Q,\omega}$ are singletons and the lattice quotient~$[e,\omega]/\equiv_{Q,\omega}$ coincides with the weak order of type $A_2$.
    On the other hand there are two facets, $I_1=\{1,3,4\}$ and $I_2=\{3,4,6\}$, which are connected by a flip but whose corresponding elements $\pi_1$ and $\pi_2$ are not related in weak order.
    \cesar{To do: Double check and expand this example in detail. Add figure.}
\end{example}

@@

\begin{conjecture}
\label{conj:latticeQuotientSortingAlternating}
For a non-empty subword complex $\subwordComplex(Q,\omega)$ where~$Q$ is sorting and alternating, the Hasse diagram of the lattice quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between acyclic facets of~$\subwordAcyclicFacets(Q, \omega)$.
\end{conjecture}

@@

\subsection{The sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}
\vincent{I have not checked this part yet.}

>>>>

\thanks{NC \& VP were partially supported by the French project CHARMS (ANR~19\,CE40\,0017), and by the French\,--\,Austrian project PAGCAP (ANR~21\,CE48\,0020 \& FWF I 5788).}

@@

% theorems
\newtheorem{theorem}{Theorem}[section]
\newtheorem{theoremA}{Theorem}
\renewcommand{\thetheoremA}{\Alph{theoremA}}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\crefname{conjecture}{Conjecture}{Conjectures}
\newtheorem{conjectureA}{Conjecture}
\renewcommand{\theconjectureA}{\Alph{conjectureA}}
\crefname{conjectureA}{Conjecture}{Conjectures}

@@

\maketitle

\tableofcontents

@@

\section{Introduction}
\label{sec:introduction}

@@

This paper develops a similar framework for acyclic pipe dreams.
Pipe dreams were introduced by N.~Bergeron and S.~Billey in~\cite{BergeronBilley} to compute Schubert polynomials and later revisited in the context of Gr\"obner geometry by A.~Knutson and E.~Miller~\cite{KnutsonMiller-GroebnerGeometry}, who coined the name \emph{pipe dreams}.
They have important connections and applications to various areas related to Schubert calculus and Schubert varieties~\cite{LascouxSchutzenberger-PolynomesSchubert, LascouxSchutzenberger-SchubertLittlewoudRichardson}. 
A pipe dream is an arrangement of pipes in the triangular shape, each entering along the vertical side and exiting along the horizontal side (see \cref{fig:pipeDreams}).
They are grouped according to their exiting permutation, given by the order in which the pipes appear along the horizontal axis.
The linear extensions of a pipe dream are the permutations of its pipes such that for each contact the northwest pipe appears before the southeast pipe in the permutation.
The pipe dreams with at least one linear extension are called acyclic and naturally appear in the study of brick polytopes~\cite{PilaudSantos-brickPolytope}.
A flip in a pipe dream exchanges a contact with a crossing between two pipes (see \cref{fig:pipeDreams}), and the flip is increasing when the contact is southwest of the crossing involved in the flip.
A brief recollection on pipe dreams is given in \cref{sec:preliminaries}.

\begin{figure}[t]
	\centerline{
		%\input{pipeDreams}
		\includegraphics[scale=.9]{pipeDreams}
	}
	\caption{Two pipe dreams of~$\pipeDreams(1365724)$ connected by an increasing flip (exchanging a contact with the crossing on the two red pipes~$3$ and~$4$).}
	\label{fig:pipeDreams}
\end{figure}

@@

\begin{theoremA}
\label{thm:A}
For any permutation~$\omega$, the Hasse diagram of the lattice quotient of the interval~$[e,\omega]$ of the weak order by the pipe dream congruence of~$\omega$ is isomorphic to the increasing flip graph on acyclic pipe dreams with exiting permutation~$\omega$.
\end{theoremA}

@@

We then explore in \cref{sec:furtherTopics} some natural further topics on pipe dreams.
We describe two algorithms to compute the unique acyclic pipe dream whose linear extensions contain a given permutation generalizing the binary tree insertion map on permutations (\cref{subsec:sweepingAlgorithm,subsec:insertionAlgorithm}), we describe the pipe dream congruence as the transitive closure of a local rewriting rule generalizing that of the sylvester congruence (\cref{subsec:rewritingRule}), and we describe a natural commutative diagram of lattice morphism generalizing the connection between the recoil map and the binary tree insertion map on permutations and the canopy on binary trees (\cref{subsec:canopy}).

@@

\begin{theoremA}
\label{thm:B}
For any non-empty subword complex~$\subwordComplex(Q, \omega)$, the sets~$\linearExtensions(I)$ for all facets~$I$ of~$\subwordComplex(Q, \omega)$ are order convex and form a partition of a lower set of the weak order that contains the interval~$[e, \omega]$.
\end{theoremA}

@@

\begin{theoremA}
\label{thm:C}
If~$Q$ is sorting, then the sets~$\linearExtensions(I)$ for all facets~$I$ of~$\subwordComplex(Q, \omega)$ form a partition of the interval~$[e, \omega]$.
\end{theoremA}

In the case when~$[e, \omega]$ does not contain all sets of linear extensions, it is natural to consider the restriction of this partition to~$[e, \omega]$ by the sets~$[e, \omega] \cap \linearExtensions(\omega)$.
This defines an equivalence relation~$\equiv_{Q, \omega}$ on~$[e, \omega]$ that we call the subword complex congruence.
However, as the sets of linear extensions are not always intervals of the weak order, this equivalence relation is not always a lattice congruence.
This seems to be fixed by an additional assumption on~$Q$.
We say that~$Q$ is alternating when all non-commuting pairs $s, t\in S$ alternate within $Q$ (this notion was already considered in \cite{PilaudSantos-brickPolytope, CeballosLabbeStump}).

\begin{conjectureA}
\label{conj:A}
If~$Q$ is alternating, then the subword complex congruence~$\equiv_{Q, \omega}$ is a lattice congruence of the interval~$[e, \omega]$ of the weak order.
\end{conjectureA}

When~$Q$ is alternating, we can thus consider the quotient~$[e, \omega]/{\equiv_{Q, \omega}}$.
In contrast to the case of pipe dreams, the Hasse diagram of the quotient~$[e, \omega]/{\equiv_{Q, \omega}}$ is not always isomorphic to the increasing flip graph on acyclic facets of~$\subwordComplex(Q, \omega)$ for two reasons:
\begin{itemize}
\item First, not all acyclic facets of~$\subwordComplex(Q, \omega)$ appear as vertices of~$[e, \omega]/{\equiv_{Q, \omega}}$ (see \cref{rem:linearExtensionsPartitionSubwordComplexA}). We say that a facet is strongly acyclic if~$[e, \omega] \cap \linearExtensions(\omega) \ne \varnothing$.
\item Second, not all flips between two strongly acyclic facets of~$\subwordComplex(Q, \omega)$ define a cover relation of~$[e, \omega]/{\equiv_{Q, \omega}}$ (see \cref{exm:acyclicFlipvsWeakOrder}). We say that the flip of a position~$i$ in a facet~$I$ is external if the root~$\rootFunction{I}{i}$ directing the flip is a ray of the root configuration~$\Roots(I)$.
\end{itemize}
This leads us to the following conjecture.

\begin{conjectureA}
\label{conj:B}
If~$Q$ is alternating, then the Hasse diagram of the quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between strongly acyclic facets of~$\subwordComplex(Q, \omega)$.
\end{conjectureA}

Combining the sorting and alternating conditions of \cref{thm:B,conj:A,conj:B}, we thus obtain the following conjecture, which can be seen as the natural extension of~\cref{thm:A}.

\begin{conjectureA}
\label{conj:C}
If~$Q$ is sorting and alternating, then the Hasse diagram of the quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between acyclic facets of~$\subwordComplex(Q, \omega)$.
\end{conjectureA}

This last conjecture has a strong connection to the geometry of subword complexes given by brick polytopes~\cite{PilaudSantos-brickPolytope, PilaudStump-brickPolytope} and brick polyhedra~\cite{JahnStump}.
Namely, the extremal increasing flips between acyclic facets of~$\subwordComplex(Q, \omega)$ is precisely the bounded part (meaning forgetting the unbounded rays) of the brick polyhedron~$\brickPolyhedron(Q, \omega)$ oriented by a natural linear functional.
\cref{conj:C} thus translates geometrically to the following.

\begin{conjectureA}
\label{conj:D}
If~$Q$ is sorting and alternating, then the oriented graph of the brick polyhedron~$\brickPolyhedron(Q, \omega)$ is isomorphic to the Hasse diagram of the lattice quotient~$[e, \omega]/\equiv_{Q, \omega}$.
\end{conjectureA}

In particular, when~$\omega = \wo$, we obtain the following conjecture, extending results from~\cite{Pilaud-brickAlgebra}.

\begin{conjectureA}
\label{conj:E}
If~$Q$ is sorting and alternating, then the oriented graph of the brick polytope $\brickPolyhedron(Q, \wo)$ is isomorphic to the Hasse diagram of the lattice quotient of the weak order by~$\equiv_{Q, \omega}$.
\end{conjectureA}

@@

@@

\begin{example}
\label{exm:Tamari1}
We say that a pipe dream is \defn{reversing} if it fixes the first and last pipes and reverses the order of the remaining pipes.
In this case, it is natural to label the pipes from~$0$ to~$n+1$, so that the permutation of the pipes is~$\rho_n \eqdef [0, n, n-1, \dots, 2, 1, n+1]$.
% We denote by~$\reversingPipeDreams_n$ the reversing pipe dreams of~$\pipeDreams_n$.
As observed in~\cite{Woo, PilaudPocchiola, Pilaud-these, Stump}, the family of reversing pipe dreams belong to the Catalan families, meaning that it is counted by the famous Catalan numbers.
\cref{fig:bijection} illustrates explicit bijections between reversing pipe dreams of~$\pipeDreams(\rho_n)$, binary trees with $n$ internal nodes, and the triangulations of a convex $(n+2)$-gon.
More precisely, the map which sends a contact in row~$i$ and column~$j$ of the triangular shape to the diagonal~$[i,n+1-j]$ of the~$(n+2)$-gon provides the following correspondence: \\[.3cm]
\centerline{
\begin{tabular}{r@{$\quad\longleftrightarrow\quad$}l}
pipe dream~$P \in \reversingPipeDreams_n$ & triangulation~$P\duality$ of the $(n+2)$-gon, \\
pipe $i$ of~$P$ & triangle $i\duality$ of~$P\duality$ (with central vertex~$i$), \\
contact between pipes~$i$ and~$j$ of~$P$ & common side of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
crossing between pipes~$i$ and~$j$ of~$P$ & common bisector of triangles~$i\duality$ and~$j\duality$ of~$P\duality$, \\
contact graph of~$P$ & dual binary tree of~$P\duality$, \\
contact flips in~$P$ & diagonal flips in~$P\duality$. \\
%increasing flip poset & Tamari lattice.
\end{tabular}} \\[.3cm]
%
\begin{figure}
	\centerline{\includegraphics[scale=1.3]{dualiteTriangulation}}
	\caption{The bijection between reversing pipe dreams (left), binary trees (middle) and triangulations (right).}
	\label{fig:bijection}
\end{figure}
%
As a consequence, this map sends the increasing flip poset on reversing pipe dreams to the \defn{Tamari lattice} on binary trees.
This lattice is defined as the transitive closure of right rotations on binary trees, and is obtained as the quotient of the weak order by the \defn{sylvester congruence}.
An important point here is that all reversing pipe dreams are acyclic, since there contact graphs are (oriented) binary trees.
\end{example}

@@

\section{Further topics on pipe dreams}
\label{sec:furtherTopics}

@@

Our first algorithm to construct~$\insertion{\pi}{\omega}$ is a \defn{sweeping algorithm}, inspired by the algorithm to compute greedy pipe dreams~\cite{PilaudPocchiola, PilaudStump-ELlabelings}.
An extension of this algorithm to subword complexes will be discussed in \cref{subsec:sweepingAlgorithmSubwordComplexes}, and appeared independently in~\cite{JahnStump}.

@@

\begin{example}
\label{exm:sweepingAlgorithm}
\cref{fig:sweepingAlgorithm} illustrates the sweeping algorithm for the permutations~${\pi = 513264}$ and~${\omega = 561324}$.
The algorithm has 15 steps, but we have grouped together steps $2$ to $9$ (second arrow) and steps $12$ to $13$ (fifth arrow) as they have the same justification.
We place contacts in steps $1$, $10$, $12$, $13$ and~$15$ since~$\omega^{-1}(i) < \omega^{-1}(j)$, and in step~$14$ since~$i > j$.
We place crossings in steps $2$ to $9$ since~$i < j$, $\omega^{-1}(i) > \omega^{-1}(j)$ and we are in the column~$\omega^{-1}(j)$, and in step $11$ since~$i < j$, $\omega^{-1}(i) > \omega^{-1}(j)$ and~$\pi^{-1}(i) > \pi^{-1}(j)$.
%
\begin{figure}[b]
	\centerline{\includegraphics[scale=.9]{sweeping}}
	\caption{Sweeping algorithm for the permutations~$\pi = 513264$ and~$\omega = 561324$.} % See details in \cref{exm:sweepingAlgorithm}.}
	\label{fig:sweepingAlgorithm}
\end{figure}
\end{example}

@@

\begin{example}
\label{exm:insertionAlgorithm}
\cref{fig:insertionAlgorithm} illustrates the insertion algorithm for the permutations~${\pi = 513264}$ and~${\omega = 561324}$.
For instance, when inserting the (green) pipe~$4$ in the last step, the currently free southeast elbows are the first southeast elbow of the (blue) pipe~$3$ and the two southeast elbows of the (purple) pipe~$2$.
%
\begin{figure}[b]
	\centerline{\includegraphics[scale=.9]{insertion}}
	\caption{Insertion algorithm for the permutations~$\pi = 513264$ and~$\omega = 561324$.}
	\label{fig:insertionAlgorithm}
\end{figure}
\end{example}

@@

The objective of this section is to partially extend our results about the lattice structure of acyclic pipe dreams (\cref{sec:latticeAcyclicPipeDreams}) to the wider context of subword complexes in finite Coxeter groups.
We start with some basic preliminaries on finite Coxeter groups (\cref{subsec:finiteCoxeterGroups}) and subword complexes (\cref{subsec:subwordComplexes}).
We then define linear extensions of facets of subword complexes (\cref{subsec:linearExtensionsFacets}) and present two theorems (\cref{subsec:twoTheorems}) and five conjectures (\cref{subsec:fiveConjectures}) about them.
To conclude, we present a sweeping algorithm to construct the facet with a given linear extension (\cref{subsec:sweepingAlgorithmSubwordComplexes})

@@

We now introduce the analogue of \cref{def:linearExtensions} for subword complexes.

@@

\begin{remark}
\label{rem:linearExtensionsPartitionSubwordComplexA}
Before proving this theorem, we would like to make the following remarks about items~\eqref{item:cover} and~\eqref{item:convex}.
\begin{enumerate}
\setcounter{enumi}{1}
\item There are cases where $[e,\omega] \ne \linearExtensions(Q,\omega)$. \cesarm{add discussion}
%
\begin{figure}[h]
	\includegraphics[width=0.52\textwidth]{interval_not_contained1} \quad
	\includegraphics[width=0.43\textwidth]{interval_not_contained2}
	\caption{A facet of a subword complex~$\subwordComplex(Q, \omega)$ whose linear extensions all lie outside the interval~$[e, \omega]$.}
	\label{fig:outOfInterval}
\end{figure}
%
\setcounter{enumi}{3}
\item $\linearExtensions(I)$ is not an interval in general. See for instance the example~\cite[Figure~9]{PilaudStump-brickPolytope}. 
In that figure, the elements of the week order $\mathfrak{S}_4$ that are linear extensions of the same acyclic facet of a given subword complex are grouped together. Some of the groups in the figure are not intervals.  \\
\end{enumerate}
\end{remark}

@@

\subsection{Five conjectures on linear extensions of facets}
\label{subsec:fiveConjectures}

@@

\begin{example}
\label{exm:acyclicFlipvsWeakOrder}
Consider the type~$A_2$ Coxeter system, the word~$Q = \tau_1 \tau_2 \tau_1 \tau_2 \tau_1 \tau_2$ and the longest element~$\wo = \tau_1 \tau_2 \tau_1 = \tau_2 \tau_1 \tau_2$.
The subword complex~$\subwordComplex(Q, \wo)$ has eight facets, six of which are acyclic.
All the fibers of $\equiv_{Q, \wo}$ are singletons and the lattice quotient~$[e, \wo]/\equiv_{Q, \wo}$ coincides with the weak order of type~$A_2$.
Note that the two acyclic facets~$\{1,3,4\}$ and $\{3,4,6\}$ are connected by a flip but the corresponding classes do not form a cover relation in~$[e, \wo]/\equiv_{Q, \wo}$.
%
\begin{figure}[h]
	\begin{tikzpicture}[xscale=1.5,->]
	\node (123) at (0,-2.5) {$\{1,2,3\}$};
	\node (134) at (-1,-1.5) {$\{1,3,4\}$};
	\node (126) at (0.3,-.5) {$\{1,2,6\}$};
	\node (236) at (1,-1.5) {$\{2,3,6\}$};
	\node (145) at (-1,1.5) {$\{1,4,5\}$};
	\node (156) at (-0.3,.5) {$\{1,5,6\}$};
	\node (346) at (1,1.5) {$\{3,4,6\}$};
	\node (456) at (0,2.5) {$\{4,5,6\}$};
	\draw (123)--(134);
	\draw (123)--(126);
	\draw (123)--(236);
	\draw (134)--(145);
	\draw (134)--(346);
	\draw (126)--(156);
	\draw (126)--(236);
	\draw (236)--(346);
	\draw (145)--(456);
	\draw (145)--(156);
	\draw (156)--(456);
	\draw (346)--(456);
	\end{tikzpicture}
	\qquad
	\begin{tikzpicture}[xscale=1.5,->]
	\node (123) at (0,-2.5) {$\{1,2,3\}$};
	\node (134) at (-1,-1.5) {$\{1,3,4\}$};
	\node (236) at (1,-1.5) {$\{2,3,6\}$};
	\node (145) at (-1,1.5) {$\{1,4,5\}$};
	\node (346) at (1,1.5) {$\{3,4,6\}$};
	\node (456) at (0,2.5) {$\{4,5,6\}$};
	\draw (123)--(134);
	\draw (123)--(236);
	\draw (134)--(145);
	\draw (134)--(346);
	\draw (236)--(346);
	\draw (145)--(456);
	\draw (346)--(456);
	\end{tikzpicture}
	\qquad
	\begin{tikzpicture}[xscale=1.5,->]
	\node (123) at (0,-2.5) {$\{1,2,3\}$};
	\node (134) at (-1,-1.5) {$\{1,3,4\}$};
	\node (236) at (1,-1.5) {$\{2,3,6\}$};
	\node (145) at (-1,1.5) {$\{1,4,5\}$};
	\node (346) at (1,1.5) {$\{3,4,6\}$};
	\node (456) at (0,2.5) {$\{4,5,6\}$};
	\draw (123)--(134);
	\draw (123)--(236);
	\draw (134)--(145);
	\draw (236)--(346);
	\draw (145)--(456);
	\draw (346)--(456);
	\end{tikzpicture}
	\caption{The increasing flip graph on~$\subwordComplex(Q, \wo)$ (left), its restriction to acyclic facets (middle) and the Hasse diagram of the quotient~$[e, \wo]/\equiv_{Q, \wo}$ where each class is labeled by its corresponding facet of~$\subwordComplex(Q, \wo)$ (right), for the subword complex~$\subwordComplex(Q, \wo)$ where~$Q = \tau_1 \tau_2 \tau_1 \tau_2 \tau_1 \tau_2$ and~$\wo = \tau_1 \tau_2 \tau_1 = \tau_2 \tau_1 \tau_2$. Note that the increasing flip~$\{1,3,4\} \to \{3,4,6\}$ is not a cover relation of~$[e, \wo]/\equiv_{Q, \wo}$.}
	\label{fig:acyclicFlipvsWeakOrder}
\end{figure}
\end{example}

@@

\begin{conjecture}
\label{conj:latticeQuotientSortingAlternating}
If~$Q$ is sorting and alternating, then the Hasse diagram of the lattice quotient~$[e, \omega]/\equiv_{Q, \omega}$ is isomorphic to the graph of extremal increasing flips between acyclic facets of~$\subwordAcyclicFacets(Q, \omega)$.
\end{conjecture}

@@

\subsection{The sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}
\vincent{I have not checked this part yet.}
\vincent{Here, we must properly cite \cite{JahnStump} and say that we obtained this algorithm independently.}
